# Peer-graded Assignment: Course Project 1
## Reproducible research

Package Info
```{r}
library(dplyr)
library(knitr)
library(ggplot2)
```

### Code for reading in the dataset and/or processing the data
```{r}
data<-read.csv("activity.csv",header=T)
summary(data)
#Remove NA data
stepsdata<-na.omit(data)
summary(stepsdata)
#Calculate total number of steps per day
by_day <- group_by(stepsdata, date)
steps_by_day <- summarise(by_day, total = sum(steps))
```
### Histogram of the total number of steps taken each day
```{r}
hist(steps_by_day$total, main="Histogram of total number of steps per day",xlab="Total number of steps in a day")
```

### Mean and median number of steps taken each day

```{r}
summary(steps_by_day)
```

### Time series plot of the average number of steps taken
Preprocessing data for plot
```{r}
steps_by_interval <- aggregate(steps ~ interval, stepsdata, mean)
# create a time series plot 
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```
 
### The 5-minute interval that, on average, contains the maximum number of steps
Find row with max of steps
```{r}
max_steps_row <- which.max(steps_by_interval$steps)
```
Find interval with this max
```{r}
steps_by_interval[max_steps_row, ]
```
### Code to describe and show a strategy for imputing missing data
```{r}
sum(is.na(data))
data_imputed <- data
for (i in 1:nrow(data_imputed))
  {
  if (is.na(data_imputed$steps[i]))
    {
      interval_value <- data_imputed$interval[i]
      steps_value <- steps_by_interval[steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```
### Histogram of the total number of steps taken each day after missing values are imputed
```{r}
  hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
# get mean and median of imputed data
mean(df_imputed_steps_by_day$steps)
median(df_imputed_steps_by_day$steps)
mean(steps_by_day$total)
median(steps_by_day$total)
```
### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
# convert type_of_day from character to factor
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)

# calculate average steps by interval across all days
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)

#create a plot
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```

